---
title: 'MY560: Reproducible Pipelines Code Walkthrough -- Tier 3'
author: "Daniel de Kadt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tier 3: Functional Programming

Let's start with a very brief introduction to the functional programming paradigm. The idea here is that everything we do in an `.R` or `.Rmd` script should be performed via functions. To a degree you are used to this already, as pretty much any package you use delivers its functionality through functions. The idea is to move our functions up one level -- everything we execute should be a contained, scoped, function that does one thing -- reads data, cleans data, merges data, runs an analysis, makes a plot, etc. 

Let's start with something really simple:

```{r add up non-functional}
# let's start with a clear workspace: 
rm(list = ls())

# inspect the elements in our workspace: 
ls()

# now run a few lines of code that produce an output: 
x <- 10
y <- 5
x + y

# inspect the elements in our workspace again. what do we find now?
ls()
```

What's happened here? `R` has saved our new objects -- x and y -- to the workspace. Let's see what happens when we use a functional approach instead: 

```{r add up functional}
# let's clear our workspace one more time: 
rm(list = ls())

# now turn out simple operation into a function
add_up <- function(x, y){
  x + y
}

# and run it
add_up(10, 5)

# check the workspace:
ls()

# clear it again:
rm(list = ls())
```

This is somewhat unintuitive at first blush, because we have moved from three lines of code to five lines of code with the same result. 

But we have gained three valuable things: First, we now have a function that, for a given set of inputs, **always** produces the exact same output (think carefully about this -- is it true of every function?). Second, once we create the function we have a way to add two scalars that we can **reuse** for any pair of scalars (and, in fact, for vectors) -- this is the basis for the modularity that comes next. Third, running our function **has no effect** on our workspace.

## Tier 3: Modular Programming

Now that we are working with functions, we can see how they support a **modular** approach to programming. That is, each function, or collection of functions (which can themselves be nested in a function), represents a module of code. Those modules are flexible in that we can change the inputs -- which changes the outputs -- but if the functions are well written we should **know** the singular mapping of inputs to outputs. 

Let's see how our very simple add_up function can work in a modular fashion: 

```{r add up modular}
source("./R/functions/function_add_up.R")
add_up(10, 5)
add_up(9, 2)
add_up(c(10,55,100), c(9,2,1))

pairs <- data.frame(x = rnorm(100,0,1), y = rnorm(100,0,1), sum = NA)

for(i in 1:nrow(pairs)){
  pairs$sum[i] <- add_up(pairs$x[i], pairs$y[i])
}

head(pairs)

# clean up the workspace:
rm(list = ls())
```

By combining functions and a modular approach we are moving toward **DRY** (Don't Repeat Yourself) code. 

## Tier 3: Building a Functional, Modular Workflow

Let's play around with the very simple workflow that we developed earlier. First, we've moved it all into a single chunk below, but it's basically the same code. Remember, we already satisfied Tier 1 (and 2) for the most part. Note that the code is currently pretty DRY, but imagine what would happen if we wanted to run a **different** analysis... Say we wanted to study how `Happiness_score` and `Life_expectancy` correlate, instead of `GDP_pc`. With our current script we would have to get **WET** -- we might copy and paste our `ggplot()` code and change the variables. Things would get even worse if we customised the code with some tweaks to the `ggplot` theme -- we'd get really WET!

```{r script workflow}
# some setup: a cheeky little bit of code to check and install packages
need <- c("tidyverse","stargazer") # list packages needed
have <- need %in% rownames(installed.packages()) # checks packages you have
if(any(!have)) install.packages(need[!have]) # install missing packages
invisible(lapply(need, library, character.only=T)) # load needed packages
 
# some more setup: pathing and folder structure
table_dir <- "./output/tables/"
figure_dir <- "./output/figures/"
data_folder <- "./data/raw/"
  
# let's start by ingesting the data
file <- "WHR_2017.csv"
data <- read_csv(paste0(data_folder,file))
head(data)

# now we will do a little data munging to two variables of interest. we will standardize these two variables. 
variables <- c("Happiness_score", "GDP_pc")

scaled_data <- data %>%
  mutate_at(variables, scale)

# let's just inspect the data, and check that it actually changed...
head(scaled_data)

data$Happiness_score[1] == scaled_data$Happiness_score[1] # should be FALSE

# next we will make a very simple two-way scatter of these two scaled. let's save the output to our output folder.
plot <- ggplot(scaled_data) +
          geom_point(aes(x = GDP_pc, y = Happiness_score)) +
          theme_minimal()
ggsave(plot, file = paste0(figure_dir,"plot_gdp_pc.jpg"))

# finally, let's run a simple regression of these two variables, and save the output as a latex table.
lm(Happiness_score ~ GDP_pc, data = data)  |>
  stargazer(out = paste0(table_dir,"table_gdp_pc.tex"))

# clean up the workspace again:
rm(list = ls())
```

Because we want to stay DRY, we're going to take a simple but important step. We are going to convert all of these to functions that can take variable arguments:

```{r functions workflow}
# our package function
package_check <- function(need = c()){
  have <- need %in% rownames(installed.packages()) # checks packages you have
  if(any(!have)) install.packages(need[!have]) # install missing packages
  invisible(lapply(need, library, character.only=T))
}

# our function to ingest the data -- this one is really easy!
ingest_data <- function(file) {
  read_csv(file)
}

# our function to scale the data -- the code changes slightly here to allow variables to be passed through
scale_data <- function(foo, variables) {
  foo_scaled <- foo %>% 
    mutate_at(variables, ~(scale(.) %>% as.vector))
  return(foo_scaled)
}

# our function to make a simple plot -- here we need to use get() to pass variable strings through
simple_plot <- function(foo, yvar, xvar, outdir, fname) {
  plot <- ggplot(foo) +
            geom_point(aes(x = get(xvar), y = get(yvar))) +
            theme_minimal()
  ggsave(plot, file = paste0(outdir, fname))
}

# our function to run a simple regression -- here we need to use get() to pass variable strings through
simple_reg <- function(foo, yvar, xvar, outdir, fname) {
  model <- lm(get(yvar) ~ get(xvar), data = foo) %>% 
    stargazer(out = paste0(outdir, filename))
}

# let's run our program, defining a few macro variables at the top feed into the various functions:
table_dir <- "./output/tables/"
figure_dir <- "./output/figures/"
package_check(c("tidyverse","stargazer"))
yvar = "Happiness_score"
xvar = "GDP_pc"

data <- ingest_data("./data/raw/WHR_2017.csv")
data_scaled <- scale_data(data, c("Happiness_score", "GDP_pc"))
simple_plot(data_scaled, yvar = yvar, xvar = xvar, outdir = figure_dir, fname = "plot_gdp_pc_function.jpg") # diff names just for show
simple_plot(data_scaled, yvar = yvar, xvar = xvar, outdir = table_dir, fname = "table_gdp_pc_function.tex")

# clean up the workspace again:
rm(list = ls())
```

This is much neater than the stream-of-consciousness script that we started with. Every step is clearly demarcated, scoped, and contained. Let's go one step further and move toward modularity. First, we are going to create a new file called `functions.R` in our functions folder. In it we will include the code we just developed above (up until the "let's run our program" line). We are then going to `source()` that file, which will generate our function objects in the workspace, and then create a new function out of those functions which executes our program. 

```{r modular functions workflow}
source("./R/functions/functions.R")

our_program <- function(table_dir = "./output/tables/", 
                        figure_dir = "./output/figures/", 
                        file = "./data/raw/WHR_2017.csv", 
                        pkgs = c("tidyverse","stargazer")){

  package_check(pkgs)
  
  yvar = "Happiness_score"
  xvar = "GDP_pc"
  
  data <- ingest_data(file)
  data_scaled <- scale_data(data, c("Happiness_score", "GDP_pc"))
  simple_plot(data_scaled, yvar = yvar, xvar = xvar, outdir = figure_dir, fname = "plot_gdp_pc_function.jpg")
  simple_reg(data_scaled, yvar = yvar, xvar = xvar, outdir = table_dir, fname = "table_gdp_pc_function.tex")
}

our_program()

# look how clean our workspace is... not bad
ls()
```

What we have here is now pretty powerful. In a few lines of top-level code we could vary the cleaning of our data, produce lots of varied analyses and plots, save them in a strong data structure. If we want to make changes to our program we know precisely where to look in the codebase. We would implement those changes, re-source, and then re-run our program. 

We could now save this script as something like `runner.R`, have it in the top directory, and be done with our pipeline... but there are some things missing, perhaps.