---
title: 'MY560: Reproducible Pipelines Code Walkthrough'
author: "Daniel de Kadt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tier 5: Groundhog for Package Dependencies

The best thing about `R` is all the packages. This is also the worst thing about `R` in terms of reproducibility, because packages change over time. Here we will walk through using `groundhog` for package version freezing. The primary alternative for this is `renv`, which is a more sophisticated but also heavier package. For most academic uses I think `groundhog` is sufficient, while `renv` is useful for more intense applications (and plays nicely with `targets` below), and for those who like using R projects (which we aren't doing here). You can see more about groundhog [here](https://groundhogr.com/), and `renv` [here](https://rstudio.github.io/renv/articles/renv.html).

```{r groundhog}

library(groundhog)

groundhog.date <- Sys.Date()-7
pkgs <- c("tidyverse", "targets")
groundhog::groundhog.library(pkgs, groundhog.date)

```

What's going on in the above code chunk? Essentially, we specify a date (this date can evolve as we work on the project, but should be locked/hard-coded once our project is 'complete'). We then specify particular packages. And then we ask groundhog to load the date-specific version of those packages. 

## Tier 5: Targets for Dependency Detection

We're now going to experiment with an automated workflow with understood dependencies. We will use the package `targets` to do so. The idea here is that we will build a directed acyclic graph (DAG) that represents the various objects (function!) in our workflow, and how they are connected.

```{r targets}
use_targets()
tar_manifest()
tar_visnetwork()

# now we execute the files, and look at our results
tar_make()
tar_read(model_life_exp)
tar_read(model_gdp_pc)
tar_read(plot_life_exp)
tar_read(plot_gdp_pc)

# let's re-inspect our workflow: 
tar_visnetwork()

# now let's go and edit our functions... then check the network... then re-run
tar_visnetwork()
tar_make()
```

One thing to note: I'm not 100% sure at this point how `groundhog` and `targets` interact. If you want to use package freezing and `targets`, it seems that `renv` is well supported -- there is a `tar_renv()` function in `targets`. 



